Case study validations of automatic bullet matching\tnoteref{t1}\tnoteref{t2}

### 2019-04-29 index.Rnw
fix typos and image sizing issues with new template

\begin{abstract}
Recent advances in microscopy have made it possible to collect 3D topographic data, enabling virtual comparisons based on the collected 3D data next to traditional comparison microscopy. Automatic matching algorithms have been introduced for various scenarios, such as matching cartridge cases (Tai and Eddy 2018) or matching bullet striae (Hare et al. 2017b,  Chu et al 2013,  De Kinder and Bonfanti 1999). One key aspect of validating automatic matching algorithms is to evaluate the performance of the algorithm on external tests. Here, we are presenting a discussion of the performance of the matching algorithm (Hare et al. 2017b) in three studies. We are considering matching performance based on the Random forest score, cross correlation, and consecutive matching striae (CMS) at the land-to-land level and, using Sequential Average Maxima scores, also at the bullet-to bullet level. Cross correlation and Random Forest scores both result in perfect discrimination of same-source and different-source bullets. At the land-to-land level, discrimination (based on area under the curve, AUC) is excellent ($ge$ 0.90).
\end{abstract}

@@

<<sam-sketch, out.width='\\textwidth', fig.width=12, fig.height = 2.5, fig.cap="Sketch of all six land-to-land sequences between two bullets with six lands.">>=
df2 <- data.frame(expand.grid(x = 1:6, y = 1:6))
df2$case1 <- df2$x == ((df2$y + 5) %% 6) + 1
df2$case2 <- df2$x == ((df2$y + 0) %% 6) + 1
df2$case3 <- df2$x == ((df2$y + 1) %% 6) + 1
df2$case4 <- df2$x == ((df2$y + 2) %% 6) + 1
df2$case5 <- df2$x == ((df2$y + 3) %% 6) + 1
df2$case6 <- df2$x == ((df2$y + 4) %% 6) + 1

@@

<<h44, echo=FALSE, fig.width = 10, fig.height=5, fig.cap="Hamby set 44: overview of all bullet-to-bullet matches between all pairs of questioned bullets (y-axis) and known test fires from ten barrels (x axis).">>=
h44features <- read.csv("data/h44-features.csv")
h44features <- h44features %>% mutate(
  lys = overlap*signature_length*1000/.645,
  cms = cms2 *lys*.645/1000
)
# tank rash lands:
#h44features <- h44features %>% mutate(
#  rfscore = replace(rfscore, (bullet1 == "I" & land1 == 6) | (bullet2 == #"I" & land2 == 6), NA)
#)
h44features <- h44features %>% mutate(
  rfscore = replace(rfscore, (barrel1 == 8 & bullet1 == "1" & land1 == 6) | (barrel2 == 8 & bullet2 == "1" & land2 == 6), NA)
)

@@

<<pd, echo=FALSE, fig.width = 9, fig.height=4, fig.cap="Phoenix study: overview of all bullet-to-bullet matches between all pairs of questioned bullets (y-axis) and known test fires from ten barrels (x axis). The order of the bullets on the y axis is determined by matching barrel.">>=
features <- read.csv("data/pd-features.csv.gz")
features <- features %>% mutate(
  lys = overlap*signature_length*1000/.645,
  cms = cms2 *lys*.645/1000
)
# change labels for land 3 and land 4 of bullet B3 in P7
features <- features %>% mutate(
  first = as.character(first),
  first = replace(first, first=="Gun 1-P7/B3/L4.dat", "XXX"),
  first = replace(first, first=="Gun 1-P7/B3/L3.dat", "Gun 1-P7/B3/L4.dat"),
  first = replace(first, first=="XXX", "Gun 1-P7/B3/L3.dat")
)

@@

<<hou, fig.width=9, fig.height = 3, fig.cap="Overview of matching scores for all pairs of questioned bullets to known bullets from five barrels (on the left) and questioned bullets to themselves (tiles on the right).", fig.subcap=c("Set 1", "Set 2", "Set 3"), fig.align="middle", fig.ncol=1>>=
features <- read.csv("data/fsi-features.csv.gz")
features <- features %>% mutate(
  lys = overlap*signature_length*1000/.645,
  cms = cms2 *lys*.645/1000
)

>>>>

\begin{abstract}
Recent advances in microscopy have made it possible to collect 3D topographic data, enabling virtual comparisons based on the collected 3D data next to traditional comparison microscopy. Automatic matching algorithms have been introduced for various scenarios, such as matching cartridge cases (Tai and Eddy 2018) or matching bullet striae (Hare et al. 2017b,  Chu et al 2013,  De Kinder and Bonfanti 1999). One key aspect of validating automatic matching algorithms is to evaluate the performance of the algorithm on external tests. Here, we are presenting a discussion of the performance of the matching algorithm (Hare et al. 2017b) in three studies. We are considering matching performance based on the Random forest score, cross correlation, and consecutive matching striae (CMS) at the land-to-land level and, using Sequential Average Maxima scores, also at the bullet-to bullet level. Cross correlation and Random Forest scores both result in perfect discrimination of same-source and different-source bullets. At the land-to-land level, discrimination (based on area under the curve, AUC) is excellent ($\ge$ 0.90).
\end{abstract}

@@

<<sam-sketch, out.width='\\textwidth', fig.width=8, fig.height = 3, fig.cap="Sketch of all six land-to-land sequences between two bullets with six lands.">>=
df2 <- data.frame(expand.grid(x = 1:6, y = 1:6))
df2$case1 <- df2$x == ((df2$y + 5) %% 6) + 1
df2$case2 <- df2$x == ((df2$y + 0) %% 6) + 1
df2$case3 <- df2$x == ((df2$y + 1) %% 6) + 1
df2$case4 <- df2$x == ((df2$y + 2) %% 6) + 1
df2$case5 <- df2$x == ((df2$y + 3) %% 6) + 1
df2$case6 <- df2$x == ((df2$y + 4) %% 6) + 1

@@

<<h44, echo=FALSE, fig.width = 8, fig.height=4, fig.cap="Hamby set 44: overview of all bullet-to-bullet matches between all pairs of questioned bullets (y-axis) and known test fires from ten barrels (x axis).">>=
h44features <- read.csv("data/h44-features.csv")
h44features <- h44features %>% mutate(
  lys = overlap*signature_length*1000/.645,
  cms = cms2 *lys*.645/1000
)
# tank rash lands:
#h44features <- h44features %>% mutate(
#  rfscore = replace(rfscore, (bullet1 == "I" & land1 == 6) | (bullet2 == #"I" & land2 == 6), NA)
#)
h44features <- h44features %>% mutate(
  rfscore = replace(rfscore, (barrel1 == 8 & bullet1 == "1" & land1 == 6) | (barrel2 == 8 & bullet2 == "1" & land2 == 6), NA)
)

@@

<<pd, echo=FALSE, fig.width = 8, fig.height=3.56, fig.cap="Phoenix study: overview of all bullet-to-bullet matches between all pairs of questioned bullets (y-axis) and known test fires from ten barrels (x axis). The order of the bullets on the y axis is determined by matching barrel.">>=
features <- read.csv("data/pd-features.csv.gz")
features <- features %>% mutate(
  lys = overlap*signature_length*1000/.645,
  cms = cms2 *lys*.645/1000
)
# change labels for land 3 and land 4 of bullet B3 in P7
features <- features %>% mutate(
  first = as.character(first),
  first = replace(first, first=="Gun 1-P7/B3/L4.dat", "XXX"),
  first = replace(first, first=="Gun 1-P7/B3/L3.dat", "Gun 1-P7/B3/L4.dat"),
  first = replace(first, first=="XXX", "Gun 1-P7/B3/L3.dat")
)

@@

<<hou, fig.width=8, fig.height = 2.67, fig.cap="Overview of matching scores for all pairs of questioned bullets to known bullets from five barrels (on the left) and questioned bullets to themselves (tiles on the right).", fig.subcap=c("Set 1", "Set 2", "Set 3"), fig.align="middle", fig.ncol=1>>=
features <- read.csv("data/fsi-features.csv.gz")
features <- features %>% mutate(
  lys = overlap*signature_length*1000/.645,
  cms = cms2 *lys*.645/1000
)
